---
title: "assignment1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

### Read data

Loading
```{r}
dat <- read.csv("activity.csv")
```

and building summary of steps per day
```{r}
dd <- dat %>% group_by(date) %>% summarise(ds = sum(steps, na.rm = TRUE))
```

### Historgram and mean total number of steps

Histogram
```{r histogram}
hist(dd$ds, breaks=10, xlab = "steps per day", main="activity")
```
and median, mean
```{r}
ave <- as.integer(round(mean(dd$ds, na.rm = TRUE), 0))
med <- median(dd$ds, na.rm = TRUE)
```

Mean number of daily steps are `r ave`, and median daily steps are `r med`

### Daily activity pattern

5-minute interval, on average across all the days in the dataset, containing the maximum number of steps:
```{r}
di <- dat %>% group_by(interval) %>% summarise(ave = mean(steps, na.rm = TRUE))
ii <- di$interval[which.max(di$ave)]
sm <- round(max(di$ave),0)
```
Interval number `r ii` is the one with the most average steps over all days (`r sm` steps)

Plotting time series of activity
```{r ts}
plot(di$interval, di$ave, type="l")
```


### Imputing missing values

Number of missing values:
```{r}
vm <- is.na(dat$steps)
vl <- length(vm)	# number of data records
vs <- sum(vm)		# number of records with missing steps
vr <- round(100 * vs / vl, 0) 		# missing values in %
```
Out of a total of `r vl` records, `r vs` are missing, so a ratio of `r vr`%

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. Good, it is late and I am tired. Replacing NA with mean.
```{r}
d2 <- dat
d2$steps[vm] <- round(mean(dat$steps, na.rm = TRUE), 0) 
```

and building summary of steps per day
```{r}
dd <- d2 %>% group_by(date) %>% summarise(ds = sum(steps, na.rm = TRUE))
```

and histogram
```{r histogram_imputed}
hist(dd$ds, breaks=10, xlab = "steps per day", main="activity")
```

and median, mean
```{r}
ave2 <- as.integer(round(mean(dd$ds, na.rm = TRUE), 0))
med2 <- as.integer(median(dd$ds, na.rm = TRUE))

# percent difference old to new ratio
1-ave/ave2

# and as a reminder, ratio of missing values
print(vr/100)
```

New, after imputation, mean number of daily steps are `r ave2`, and median daily steps are `r med2`.



### Weekdays and weekends

converting to real dates and generate weekday, weekend factor
```{r}
dates <- strptime(d2$date, format="%Y-%m-%d")
wdays <- weekdays(dates)
we <- wdays %in% c("Saturday",  "Sunday")
d2$ww <- factor(we, labels = c("weekday", "weekend"))
```

5-minute interval, on average across all the days but separating out weekends from weekdays:
```{r}
di <- d2 %>% group_by(ww, interval) %>% summarise(ave = mean(steps, na.rm = TRUE))
```



and plotting the time series of the activities
```{r ts2}
library(lattice)
xyplot( ave ~ interval  | ww, data = di, type = "l", ylab="average per interval")
```
